### 2024-11-30 paper/paper.qmd
world bank refernce added

---
title: "Perceptions of Corruption in Nigeria"
subtitle: "Rural Nigerians and Less Educated  Groups Report Higher Corruption"
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at:https://github.com/fatimahsy/Nigeria_Democracy-/tree/main"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false
#| label: setup

#install.packages("modelsummary")
#install.packages("rstanarm")
#install.packages("dplyr")
#install.packages("here")
#install.packages("arrow")
library(modelsummary)
library(rstanarm)
library(dplyr)
library(here)
library(arrow)
library(bayesplot)
library(ggplot2)
library(arrow)
library(here)

library(patchwork)

# Load data
cleaned_data <- read_parquet(here::here("data/02-analysis_data/cleaned_nigeriasurvey.parquet"))
bayesian_model <- readRDS(file = here::here ("models/bayesian_model.rds"))
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as
follows. @sec-data....

We use the statistical programming language R [@citeR]....

# Data {#sec-data}

## Measurement

For this analysis, i use the Afrobarometer Rouund 9, Nigeria, 2023
survey. Afrobarometer conducts surveys that are nationally
representative across African nations to capture public perceptions on
topics like governance, democracy, and economic issues. The Nigerian
Round 9 surveey dataset contains responses from 1,556 people and is
refective of the diverse demographic statistics across gender, region,
and residence type(urban or rural). The data is freely available on the
Afrobarometer website[@afrobarometer2023] and was collected through a
multi-stage, stratified sampling method. This is to ensure
representativeness and enables analysis of different demographics and
socioeconomic indicators.

Our primary measurement task is to capture public opininon towards
economic, political and socioeconomic issues and translate it into
actionable intell. The Afrobarometer survey translates these real world
perceptions into quantifiable data using survey instruments.The process
begins with identifying relevant phenomena in the Nigerian
socio-political and economic context, such as public attitudes toward
democracy, demand for governance accountability, and perceived economic
conditions. These phenomena are operationalized into survey questions
that are clear, culturally sensitive, and tailored to elicit meaningful
responses.

For example:

-   Perceptions of Democracy: A phenomenon such as the demand for
    democracy is operationalized by questions asking respondents whether
    they prefer democracy over other forms of governance, and whether
    they believe Nigeria is a democratic nation. These questions are
    grounded in observable public discourse, historical trends, and
    policy outcomes.
-   Economic Prosperity: Attitudes toward economic conditions are
    measured by asking individuals to rate the current state of the
    economy and their own living standards. These subjective evaluations
    stem from tangible experiences like inflation, unemployment, or
    changes in public service delivery.

Each survey question acts as a data proxy for these complex real-world
dynamics, reducing them to quantifiable variables while maintaining
their conceptual integrity.

This dataset captures a unique moment in Nigeria's socio-political
landscape, reflecting how citizens perceive democracy and governance in
the context of ongoing economic and political challenges. Alternative
datasets, such as Nigeria’s national statistics or World Bank survey
data, could provide economic indicators but lack the granularity of
Afrobarometer’s public opinion data.

## Variables of Interest
```{r}
#| label: fig-urban-rural-distribution
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: " Urban vs Rural Respondent Distribution. This bar chart compares the number of respondents residing in urban versus rural areas."

# Plot: Urban vs Rural Respondent Distribution
cleaned_data %>%
  count(URBRUR) %>%
  ggplot(aes(x = URBRUR, y = n, fill = URBRUR)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Urban vs Rural Respondent Distribution",
    x = "Respondent Location",
    y = "Number of Respondents"
  ) +
  theme_minimal() +
  scale_fill_manual(
    values = c("Urban Residents" = "blue", "Rural Residents" = "gray")
  ) +
  guides(fill = "none")

```

@fig-urban-rural-distribution illustrates the distribution of survey
respondents based on their residential location, categorized as urban or
rural. From this, we see that the sample has more rural residents
compared to urban residents. Specifically, the Afrobarometer survey for
Nigeria, conducted in 2022, reports that approximately 53% of Nigeria's
population resides in rural areas, this is is consistent with national
demographic trends [@worldbank2022rural]. This rural-urban distribution
is important for helping us understand the perspectives seen in the
survey, since rural populations may have different challenges in
governance and corruption compared to urban populations. The imbalance
in representation shows that looking at geographical disparities when
interpreting results from nationwide surveys is important.

```{r}
#| label: fig-corruption-presidency
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Perceived Corruption in the Presidency. This bar chart shows the distribution of respondents' perceptions of corruption in the presidency."

# Plot: Perceived Corruption in the Presidency
cleaned_data %>%
  count(corruption_presidency) %>%
  ggplot(aes(x = corruption_presidency, y = n, fill = corruption_presidency)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Perceived Corruption in the Presidency",
    x = "Perception of Corruption",
    y = "Number of Respondents"
  ) +
  theme_minimal() +
  scale_fill_manual(
    values = c("No Corruption" = "green", 
               "Some Corruption" = "yellow", 
               "High Corruption" = "red", 
               "All Corrupt" = "purple")
  ) +
  guides(fill = "none")

```

@fig-corruption-presidency presents the distribution of respondents'
perceptions of corruption within the Nigerian presidency. The variable 
"Perceived Corruption in the Presidency" captures respondents' perceptions of 
the extent of corruption within the Nigerian presidency and its officials. This 
variable originates from Question 38A of the Afrobarometer survey, which asked:
*"How many of the following people do you think are involved in corruption, or 
haven’t you heard enough about them to say: the president and officials in his 
office?"* [@Afrobarometer2022codebook] Respondents could choose from the following values:

-   **0 (None)**: Indicates that the respondent perceives no corruption
    in the office of the presidency.
-   **1 (Some of them)**: Suggests that the respondent believes a few
    individuals in the presidency are involved in corruption.
-   **2 (Most of them)**: Reflects the perception that corruption is
    prevalent among the majority of the presidency's officials.
-   **3 (All of them)**: Suggests the respondent perceives widespread
    corruption, implicating all officials in the presidency.
-   **8 (Refused)**: Indicates the respondent declined to answer the
    question.
-   **9 (Don’t know/Haven’t heard enough)**: Suggests the respondent was
    unsure or lacked sufficient knowledge to provide an answer.
-   **-1 (Missing)**: Represents missing or unavailable responses.

For the analysis, only the substantive responses (0-3) were considered,
in order to help us look into perceptions of corruption within the presidency. 
The high frequency of responses in the "High Corruption" and "Some
Corruption" categories reflects persistent concerns about the integrity
of key officials in Nigeria, aligning with broader findings on
corruption in the region.@fig-corruption-presidency presents the distribution of 
respondents' perceptions of corruption within the Nigerian presidency. Among
these, the category "High corruption" is the most frequently reported,
followed by "Some corruption". "All corrupt" ranks third, with "No
Corruption" being the least chosen category which is no surprise given
our data is in the context of Nigeria. These proportions show
significant public concern about corruption within the presidency,
consistent with broader trends in public concern abut corruption
corruption within the presidency in Nigeria.

```{r}
#| label: fig-age-group-distribution
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Age Group Distribution. This bar chart displays the number of respondents in each age group."

# Plot: Age Group Distribution
cleaned_data %>%
  count(age_group) %>%
  ggplot(aes(x = age_group, y = n, fill = age_group)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Age Group Distribution",
    x = "Age Group",
    y = "Number of Respondents"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  guides(fill = "none")

```
@fig-age-group-distribution illustrates the distribution of survey respondents across five distinct age groups: 18-24, 25-34, 35-44, 45-54, and 55+. The data reveals that the 25-34 age group is the most represented, with over 500 respondents, followed by the 35-44 age group. The youngest cohort, aged 18-24, comprises approximately 300 respondents, while the older age groups (45-54 and 55+) are less represented in the dataset, with fewer than 200 respondents each. This distribution highlights a concentration of responses from younger demographics, particularly those aged 25-34, which may reflect broader societal trends such as population age structure or survey reach and accessibility.
```{r}
#| label: fig-trust-political-institutions
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Distribution of Trust in Political Institutions. This histogram shows how respondents rated their trust in political institutions, from low to high trust."

library(haven)

# Convert haven_labelled variables to appropriate types
cleaned_data <- cleaned_data %>%
  mutate(
    Q38A = as.numeric(as.character(Q38A)), # Convert to numeric
    Q94 = as.numeric(as.character(Q94)),   # Convert to numeric
    Q101 = as.numeric(as.character(Q101)), # Convert to numeric if needed
    URBRUR = as_factor(URBRUR),            # Convert to factor
    corruption_presidency = as_factor(corruption_presidency), # Convert to factor
    urban_rural = as_factor(urban_rural)   # Convert to factor
  )

# Plot: Trust in Political Institutions
ggplot(cleaned_data, aes(x = Q38A)) +
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") +
  labs(
    title = "Trust in Political Institutions",
    x = "Trust Level (1 = Low, 5 = High)",
    y = "Frequency"
  ) +
  theme_minimal()

```

```{r}
#| label: fig-demand-democracy
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Distribution of Demand for Democracy. This histogram illustrates respondents' levels of demand for democracy, with higher values reflecting stronger democratic preferences."

# Plot: Demand for Democracy
ggplot(cleaned_data, aes(x = Q94)) +
  geom_histogram(binwidth = 1, fill = "orange", color = "black") +
  labs(
    title = "Demand for Democracy",
    x = "Demand Level (1 = Low, 5 = High)",
    y = "Frequency"
  ) +
  theme_minimal()

```

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome
variable or update the subheading to be singular.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine
a few into one if they go together naturally.

Background details and diagnostics are included in [Appendix
-@sec-model-details].

# Model {#sec-model}

## Model set-up

A Bayesian ordinal logistic regression model is used to estimate the
relationship between various demographic factors and perceptions of
corruption in the presidency. Ordinal logistic regression is a
statistical technique used for ordered categorical outcomes to predict
the cumulative probability of being in a particular category or below,
given a set of predictor variables.

My model includes four independent demographic variables: `education`,
`gender`, `urban_rural`, and `age_group`. The dependent variable is
`corruption_presidency`, which has four ordered categories: "None,"
"Some," "Most," and "All."

The ordinal logistic regression model I used is: \begin{equation}
\log\left(\frac{P(y_i \leq j)}{P(y_i > j)}\right) = \alpha_j - \left(\beta_1 \times \text{education} + \beta_2 \times \text{gender} + \beta_3 \times \text{urban rural} + \beta_4 \times \text{age group}\right)
\end{equation}

$$
\begin{aligned}
\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
\beta_2 & \sim \mbox{Normal}(0, 2.5)\\
\beta_3 & \sim \mbox{Normal}(0, 2.5)\\
\beta_4 & \sim \mbox{Normal}(0, 2.5)\\
\alpha_j & \sim \mbox{Normal}(0, 2.5) \quad \text{for } j = 1, 2, 3\\
\pi_{ij} & \sim \mbox{Dirichlet}(\alpha = 1)
\end{aligned}
$$ where:

-   $y_i$ represents the ordinal outcome for respondent $i$, indicating
    their perception of corruption in the presidency.

-   $P(y_i \leq j)$ is the cumulative probability of respondent $i$
    falling into category $j$ or below.

-   $\alpha_j$ are the threshold parameters that define the boundaries
    between the corruption categories.

-   $\beta_1, \beta_2, \beta_3, \beta_4$ are the coefficients
    corresponding to the predictors:

    -   $\beta_1$: Education level of the respondent.
    -   $\beta_2$: Gender of the respondent (male or female).
    -   $\beta_3$: Urban or rural location of the respondent.
    -   $\beta_4$: Age group of the respondent.

-   $\pi_{ij}$ is the probability that respondent $i$ falls into
    corruption category $j$.

In this model, the priors reflect a belief that the predictors explain
about 20% of the variability in the outcome, as defined by the
$R^2(0.2)$ prior. The Normal priors with a mean of 0 and standard
deviation of 2.5 for the coefficients and threshold parameters are
relatively weakly informative, ensuring the data have a significant role
in determining the posterior distributions.

The threshold parameters ($\alpha_j$) separate the ordered categories of
the outcome variable, while the Dirichlet prior for $\pi_{ij}$ assumes
no strong prior preference for any specific corruption category.

The use of these priors provides flexibility for the model while
preventing overfitting. The posterior distributions for the coefficients
and thresholds, along with diagnostic checks for model convergence and
predictive accuracy, can be found in @sec-appendix.

### Model Assumptions

The model relies on the following assumptions:

-   Proportional Odds: The relationship between the predictors and the
    log-odds of being in a higher corruption category is constant across
    thresholds.
-   Ordinal Outcomes: The response categories are ordered, but the
    distances between them are not quantitatively defined.
-   Independence of observations: Responses from different individuals
    are assumed to be independent.

## Model Justification

The Bayesian approach was chosen because of its flexibility in
incorporating prior beliefs and its ability to provide real
quantification for uncertainty for parameter estimates. An ordinal
logistic regression model is appropriate because the outcome variable
(perceived corruption) is inherently ordered.

The predictors-education, gender, urban/rural location, and age group-
were selected based on their theoretical relevance to perceptions of
governmental corruption in Nigeria, as discussed in the data section.
For example, urban and rural individuals may influence exposure to
governance systems, while education might shape perceptions of systemic
corruption.

The modeling decisions align closely with the structure and
characteristics of the dataset to ensure interpretability and relevance.
Age was included as a categorical variable reflecting predefined groups,
allowing the model to capture non-linear relationships between age and
perceptions of corruption. Education was treated as a categorical
variable to account for differences across distinct milestones, such as
"Primary school completed" and "Post-graduate qualifications." Gender
was modeled as a binary variable ("Male" or "Female") to reflect
societal influences on corruption perceptions. Urban versus rural
context was included as a binary variable to capture potential
differences in exposure to governance and information access. While the
dataset includes regional information, province effects were excluded to
maintain focus on demographic factors and simplify the model. Finally,
the ordinal logistic regression framework was chosen to respect the
ordered nature of the outcome variable ("None," "Some," "Most," "All")
without assuming equal distances between categories. These decisions
ensure the model is both consistent with the data and well-suited to
addressing the research question.

# Results

## Corruption Perceptions Model

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Summary of the Ordinal Logistic regression model"
#| warning: false

#install.packages("broom.mixed")
#install.packages("kableExtra")

library(broom.mixed)
library(kableExtra)

# Extract summary
model_summary <- broom.mixed::tidy(bayesian_model)

# Render as a table
knitr::kable(
  model_summary,
  caption = "Summary of the Bayesian Model",
  digits = 2
) %>%
  kableExtra::kable_styling()

```

The analysis revealed distinct trends in Corruption perceptions in
different demographics. Our results are summarized in
@tbl-modelresults.This presents the estimated coefficients and standard
errors for each predictor included in the model. The dependent variable
is the perception of corruption in the presidency, categorized into
ordered levels. The table includes predictors such as education level,
gender, urban or rural location, and agre group, which are modeled to
explain variations in corruption perceptions. Each predictor's
coefficient represents its effect on the log odds of perceiving
corruption in the presidency at a specific level relative to lower
levels, holding other variables constant.

Higher education levels(e.g., university or post-graduate
qualifications) arre associated with positive effects on corruption
perceptions, indicating a greater likelihood of perceived higher levels
of corruption. Begin female shows a small positive association with
higher corruption perceptions. Residing in rural areas is negatively
associated with corruption perceptions, suggesting lower perceptions of
corruption compared to urban counterparts. Agegroups show minimal
variation in their association with corruption perceptions, with no
significant deviations observed among different age brackets.

## Predicted Probabilities

```{r}
#| echo: false
#| eval: true
#| label: fig-predicted-probabilities
#| fig-cap: "Predicted probabilities of high corruption perceptions by key predictors."

library(ggplot2)

# Simulating or using your posterior data
predicted_data <- data.frame(
  Predictor = names(coef(bayesian_model)),
  Estimate = coef(bayesian_model)
)

# Plot with adjusted x-axis labels
ggplot(predicted_data, aes(x = Predictor, y = Estimate)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Rotate and reduce size
    axis.title.x = element_text(size = 10)
  ) +
  labs(
    title = "Predicted Probabilities by Demographics",
    x = "Predictor",
    y = "Probability"
  )

```

@fig-predicted-probabilities illustrates the predicted probabilities
derived from the Bayesian ordinal logistic regression model for
different demographic predictors influencing corruption perceptions in
Nigeria. Each bar represents the estimated impact of a specific
predictor on the likelihood of perceiving higher corruption levels in
the presidency.

Higher education levels consistently show positive probabilities,
indicating that individuals with more advanced education levels are more
likely to perceive higher corruption levels. Living in rural areas has a
significantly negative probability, suggesting that rural residents are
less likely to perceive high corruption compared to urban residents.
Gender shows a moderate positive probability, with females slightly more
likely to perceive higher corruption levels than males. There seems to
be minimal variation observed across age groups, suggesting that age
does not substantially influence perceptions of corruption.

The results show that higher levels of education, particularly
university completion and post-secondary qualifications, are associated
with greater perceptions of corruption in the presidency. Urban
residents demonstrate significantly higher perceived corruption levels
compared to rural residents. Gender differences reveal that female
respondents show slightly higher corruption perceptions, while age
groups show minimal variation. These findings are shown in the
regression estimates and predicted probabilities, as visualized in
@fig-predicted-probabilities and summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The
discussion is a chance to show off what you know and what you learnt
from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what
your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior Predictive Check

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck
#| fig-cap: "Posterior Predictive Check for the Bayesian ordinal logistic regression model."

# Load the Bayesian model
bayesian_model <- readRDS(file = here::here("models/bayesian_model.rds"))

# Posterior Predictive Check
pp_check(bayesian_model) +
  theme_classic() +
  labs(
    title = "Posterior Predictive Check",
    x = "Observed Data",
    y = "Density"
  )

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-posteriorvsprior-education
#| fig-cap: "Posterior vs Prior Comparison for Education Parameters."

# Education Parameters
education_params <- c(
  "educationInformal schooling only",
  "educationSome primary schooling",
  "educationPrimary school completed",
  "educationSome secondary schooling",
  "educationSecondary school completed",
  "educationPost-secondary qualifications",
  "educationSome university",
  "educationUniversity completed",
  "educationPost-graduate"
)

posterior_vs_prior(bayesian_model, pars = education_params, group_by_parameter = TRUE) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(
    title = "Posterior vs Prior Comparison (Education Parameters)",
    x = "Parameter Value",
    y = "Density",
    color = "Distribution"
  )

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-posteriorvsprior-demographic
#| fig-cap: "Posterior vs Prior Comparison for Demographic Parameters."

# Demographic Parameters
demographic_params <- c(
  "genderFemale",
  "urban_ruralRural",
  "age_group25-34",
  "age_group35-44",
  "age_group45-54",
  "age_group55+"
)

posterior_vs_prior(bayesian_model, pars = demographic_params, group_by_parameter = TRUE) +
  theme_minimal() +
  scale_color_brewer(palette = "Set2") +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(
    title = "Posterior vs Prior Comparison (Demographic Parameters)",
    x = "Parameter Value",
    y = "Density",
    color = "Distribution"
  )

```

## Diagnostics

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-diagnostic-ppcheck
#| fig-cap: "Posterior Predictive Check for the Bayesian ordinal logistic regression model."

# Posterior Predictive Check
pp_check(bayesian_model) +
  theme_classic() +
  labs(
    title = "Posterior Predictive Check",
    x = "Observed Data",
    y = "Density"
  )

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-diagnostic-traceplot
#| fig-cap: "Trace plot for selected parameters in the Bayesian ordinal logistic regression model."

# Trace Plot
mcmc_trace(as.array(bayesian_model), pars = c("educationSome primary schooling", "urban_ruralRural")) +
  theme_minimal() +
  labs(
    title = "Trace Plot for Selected Parameters",
    x = "Iteration",
    y = "Parameter Value"
  )

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-diagnostic-rhat
#| fig-cap: "Rhat values for model convergence in the Bayesian ordinal logistic regression model."

# Rhat Plot
mcmc_rhat(rhat(bayesian_model)) +
  theme_minimal() +
  labs(
    title = "Rhat Plot",
    x = "Rhat Values",
    y = "Parameters"
  )

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-diagnostic-ess
#| fig-cap: "Effective sample size plot for parameters in the Bayesian ordinal logistic regression model."

# Effective Sample Size Plot
mcmc_neff(neff_ratio(bayesian_model)) +
  theme_minimal() +
  labs(
    title = "Effective Sample Size Plot",
    x = "Effective Sample Size Ratio",
    y = "Parameters"
  )

```

\newpage

# References