### 2024-12-03 paper/paper.qmd
prose checks

---
title: "Shaping Anti-Corruption Campaigns in Nigeria"
subtitle: "Youth Populations, Urban Areas and Moderately Educated Citizens Are Important for Anti-Corruption Campaigns "
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at:https://github.com/fatimahsy/Nigeria_Democracy-"
date: today
date-format: long
abstract: "Peceptions of corruption in the Nigerian presidency vary significantly across demogrpahic groups. Using Afrobarometer survey data and Bayesisan ordinal logistic regression, this study looks at how education, gender, age, and urban versus rural residence influence these perceptions. The results show that urban residents and yourger individuals are more likely to perceive high corruption, while rura populations and those with higher education show views that are moderate. These results show the need for targeted anti-corruption campaigns to adress demographic differences and imporve governance accountability in Nigeria"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false
#| label: setup

library(modelsummary)
library(rstanarm)
library(dplyr)
library(here)
library(arrow)
library(bayesplot)
library(ggplot2)
library(arrow)
library(here)
library(patchwork)

# Load data
cleaned_data <- read_parquet(here::here("data/02-analysis_data/cleaned_nigeriasurvey.parquet"))
bayesian_model <- readRDS(file = here::here ("models/bayesian_model.rds"))
```

# Introduction

Corruption is a very big challenge to governance and development in Nigeria as a result of its ability to undermine public trust and economic growth. As the United Nations Office on Drugs and Crime (UNODC) stated, "corruption affects the lives of ordinary Nigerian, impeding development and eroding public trust in institutions [@corruptionsurvey2024]. This paper looks into the role of demographic factors-education,gender,age, and urban versus rural residence-in shaping perceptions of corruption in the Nigerian presidency.

Primarily, this paper is aimed at understanding how different demographic groups perceive corruption in the Nigerian presidency and to identify ways thiis can help improve anti-corruption campaigns, While existing studies have largely been focused on the institutional dynamics, there seems to be a lack of demographic-specific analysis within the Nigerian context. This paper addresses this gap by making these demographics the focus. It does this by using nationally representative data from Afrobarometer [@afrobarometer2023], a leading survey initiative capturing public attitudes on governance, democracy, and corruption across Africa. A Bayesian ordinal logistic regression was used to find patterns in how education, gender, age, and geographic location influence perceptions of corruption.

Our paper shows that there are alot of disparities in corruption perceptions across demographic groups. Younger respondents and urban residents have higher perceptions of corruption, likely because of higer exposure to investigative media and political discourse. In contrast, rural populations-often reliant on patronage systems- tend to report more neutral views. Education also plays an important role since we find that secondary school graduates have the highest corruption perceptions.

Understanding these demographic differences is important for designing effective anti-corruption campaigns. Making sure outreach efforts are tailored to the unique experiences and perceptions of different demographic groups can increase their effectiveness and increase public trust. For example, youth-focused campaigns that use social media platforms and urban outreach using investigative journalism can engage the people. At the same time, localized, culturally appropriate campaigns in rural areas can address the different needs of these populations. A the UNODC emphasizes, 'targeted and effective anti-corruption policies will improve likes of Nigerian citizens" [@afrobarometer2023].

The remainder of this paper is structured as follows; @sec-data detailed the data used for this paper, @sec-model introduces the Bayesian ordinal logistic regression model used to study the relationship between our variables of interest. @sec-results presents the findings from the model, detailing how different demographic factors like education and age affect corruption perception in Nigeria. @sec-discussion examines these results, discussing their implications for anti-corruption policies, weaknesses and next steps. @sec-survey-appendix explores the details of the Afrobarometer survey. @sec-model-details shows plots for Posterior predictive checks and Diagnostics as well as their analysis. @sec-survey-appendix details the survey methodology used in the Afrobaromenter data and @sec-ethics addresses very important ethical principles and the ethical challenges of surveying sensitive topics like corruption.

# Data {#sec-data}

## Measurement

For this analysis, i use the Afrobarometer Round 9, Nigeria, 2023
survey. Afrobarometer conducts surveys that are nationally
representative across African nations to capture public perceptions on
topics like governance, democracy, and economic issues. The Nigerian
Round 9 survey dataset contains responses from 1,556 people and shows
the diverse demographic statistics across gender, region, and residence
type(urban or rural). The data is freely available on the Afrobarometer
website[@afrobarometer2023] and was collected through a multi-stage,
stratified sampling method. This is to ensure representativeness and
enables analysis of different demographics and socioeconomic indicators.

Our primary measurement task is to capture public opinion towards
economic, political and socioeconomic issues and translate it into
actionable lintel. The Afrobarometer survey translates these real world
perceptions into quantifiable data using survey instruments.The process
begins with identifying relevant phenomena in the Nigerian
socio-political and economic context, such as public attitudes toward
democracy, demand for governance accountability, and perceived economic
conditions. These phenomena are operationalized into survey questions
that are clear, culturally sensitive, and tailored to elicit meaningful
responses [@Afrobarometer2022codebook].

For example:

-   Perceptions of Democracy: A phenomenon such as the demand for
    democracy is operationalized by questions asking respondents whether
    they prefer democracy over other forms of governance, and whether
    they believe Nigeria is a democratic nation. These questions are
    grounded in observable public discourse, historical trends, and
    policy outcomes.
-   Economic Prosperity: Attitudes toward economic conditions are
    measured by asking individuals to rate the current state of the
    economy and their own living standards. These subjective evaluations
    stem from tangible experiences like inflation, unemployment, or
    changes in public service delivery.

Each survey question acts as a data proxy for these real-world
dynamics, reducing them to quantifiable variables while keeping
their conceptual integrity.This dataset shows a unique moment in Nigeria's socio-political
landscape, showing how citizens perceive democracy and governance in
the context of ongoing economic and political challenges. Alternative
datasets, such as Nigeria’s national statistics or World Bank survey
data, could provide economic indicators but do not have the granularity of
Afrobarometer’s public opinion data.

The data for this study was downloaded, cleaned, analyzed, modeled, and
visualized using R [@citeR], a statistical programming language.[@afrobarometer2023] served as the primary data source, providing importantsurvey data for analysis.***Telling Stories with Data*** [@tellingstories] was referenced for its code and methodologies. The
following packages were used:

-   **`modelsummary`** [@modelsummary]: Used to create well-formatted
    regression tables and summaries.
-   **`rstanarm`** [@rstanarm]: Facilitated Bayesian regression modeling
    using the Stan backend.
-   **`dplyr`** [@dplyr]: Streamlined data manipulation, transformation,
    and summarization processes.
-   **`here`** [@here]: Simplified the organization of file paths,
    improving project reproducibility.
-   **`arrow`** [@arrow]: Enabled efficient handling of large datasets
    with fast reading and writing capabilities.
-   **`bayesplot`** [@bayesplot]: Assisted in producing diagnostic
    checks and posterior predictive visualizations for Bayesian models.
-   **`ggplot2`** [@ggplot2]: Provided extensive tools for creating
    clear and impactful visualizations.
-   **`patchwork`** [@patchwork]: Combined multiple plots into cohesive,
    professional visuals.
-   **`knitr`** [@knitr]: Enabled dynamic report generation, integrating
    code, output, and text seamlessly.
-   **`kableExtra`** [@kableExtra]: Enhanced table styling and
    formatting for clean, professional presentations.
-   **`ggalt`** [@ggalt]: Allowed the creation of specialized plots,
    enhancing the depth of data visualizations.
-   **`haven`** [@haven]: Facilitated the import of `.sav` survey data
    from the Afrobarometer.
-   **`broom.mixed`** [@broommixed]: Streamlined the extraction and
    tidying of results from mixed models for easy analysis and
    visualization.
-   ***Telling Stories with Data*** [@tellingstories]: This book was
    referenced for its code and methodologies.

## Cleaned Data Overview

```{r}
#| label: tbl-cleaned-data-preview
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| tbl-cap: "Preview of the Cleaned Dataset"

library(knitr)
library(kableExtra)

# Create a preview of the cleaned dataset
table_preview <- cleaned_data %>% 
  select(-c(1:6)) %>%
  head(6)

# Generate a well-formatted table using kableExtra
kable(
  table_preview, 
  caption = "Preview of the Cleaned Dataset"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

@tbl-cleaned-data-preview represents the first six cleaned rows of the
dataset, showing the important variables used in the analysis. The
variables were constructed from specific survey questions in the
Afrobarometer Round 9 Nigeria Survey [@afrobarometer2023], which
captures several socioeconomic sentiments within the country.The column
corruption_presidency captures perceptions of corruption within the
office of the presidency, derived from responses to the question, “How
many of the following people do you think are involved in corruption, or
haven’t you heard enough about them to say: the president and officials
in his office?” The variable education reflects respondents' highest
level of education completed, gathered from the question, “What is the
highest level of education you have completed?” The gender variable
indicates the respondent's gender, collected through direct demographic
inquiry. The column age_group categorizes respondents into age brackets
based on the question, “How old are you?” The urban_rural variable
distinguishes between urban and rural respondents, as recorded by
interviewers based on the primary sampling unit (PSU) during data
collection [@survey_manual_2022]. Lastly, region specifies the
geographical region of each respondent, identified using a unique code
for Nigerian states. This table not only demonstrates the diversity of
the dataset but also highlights the breadth of demographic and
governance-related variables included for analysis.

```{r}
#| label: tbl-summary-statistics
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| tbl-cap: "Detailed Summary Statistics of the Cleaned Dataset"

library(dplyr)
library(knitr)
library(kableExtra)

# Ensure the age_group column is a character for correct mapping
cleaned_data <- cleaned_data %>%
  mutate(age_group = as.character(age_group))

# Function to calculate the mode (most frequent category)
calculate_mode <- function(x) {
  x <- na.omit(x)
  unique_x <- unique(x)
  unique_x[which.max(tabulate(match(x, unique_x)))]
}

# Adjust line breaks for proper rendering
line_break <- ifelse(knitr::is_html_output(), "<br>", "\n")

# Prepare the data for the transposed table
transposed_summary <- tibble(
  Statistic = c(
    "Most Frequent Age Group", 
    "Urban Percentage", 
    "Male Percentage", 
    "Corruption Counts", 
    "Education Counts"
  ),
  Value = c(
    calculate_mode(cleaned_data$age_group), 
    paste0(round(mean(cleaned_data$urban_rural == "Urban Residents", na.rm = TRUE) * 100, 2), "%"),
    paste0(round(mean(cleaned_data$gender == "Male", na.rm = TRUE) * 100, 2), "%"),
    paste0(
      "None: 61", line_break,
      "Some: 574", line_break,
      "High: 593", line_break,
      "All: 328"
    ),
    paste0(
      "Primary: 203", line_break,
      "Secondary: 173", line_break,
      "University: 102", line_break,
      "Postgraduate: 07"
    )
  )
)

# Generate a formatted table
kable(
  transposed_summary, 
  col.names = c("Statistic", "Value"),
  caption = "Detailed Summary Statistics of the Cleaned Dataset (Transposed)",
  escape = FALSE # Disables escaping for proper HTML/PDF formatting
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(2, width = "20em") # Adjust width for better multiline display

```

The cleaned dataset used in this study is summarised in
@tbl-summary-statistics, which provides important statistics to provide an
overview of the dataset's structure and characteristics. The most
frequent age group among respondents is identified, representing the
predominant demographic in terms of age. The table also highlights the
urban and male percentages, which indicate the proportion of urban
residents (43.38%) and male respondents (49.68%) in the sample.

Further, the table breaks down the perceptions of corruption into four
categories—"None," "Some," "High," and "All"—with counts for each,
offering information into respondents' views on corruption in the
presidency. Additionally, the table details the distribution of
respondents across education levels, including categories such as
"Primary," "Secondary," "University," and "Postgraduate." Together,
these summary statistics provide a very detailed snapshot of the
dataset, enabling a clear understanding of its demographic and
categorical variables.

## Corruption Perceptions Vs Variables of Interest

```{r}
#| label: fig-education-vs-corruption
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Relationship between Education Levels and corruption percetpions in Nigeria  in Nigeria"

cleaned_data %>%
  count(education, corruption_presidency) %>%
  ggplot(aes(x = corruption_presidency, y = education, fill = n)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(
    title = "Heatmap of Education and Corruption Perceptions",
    x = "Corruption Perception",
    y = "Education Level",
    fill = "Number of Respondents"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5), # Reduce title size and center it
    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels
    axis.text.y = element_text(size = 10) # Adjust y-axis label size
  )

```

This heatmap in @fig-education-vs-corruption shows the relationship
between respondents’ education levels and their perceptions of
corruption in Nigeria. The data reflects responses to the question:
*What is your highest level of education?*.The x-axis represents
corruption perception categories (All Corrupt, High Corruption, No
Corruption, and Some Corruption), while the y-axis represents education
levels, ranging from No Formal Schooling to Post-graduate. The intensity
of the color reflects the number of respondents, with darker shades
indicating higher counts.

The heatmap shows that respondents who completed secondary school or
some secondary schooling dominate the High Corruption and Some
Corruption categories. Respondents with higher education levels
(university and post-graduate) are distributed more evenly across the
corruption perception categories, with fewer concentrated in any
specific category. Conversely, respondents with no formal schooling or
only informal education are sparsely represented across all
categories.There is a notable concentration of corruption perceptions
among respondents with mid-level education (secondary schooling), while
those with higher education tend to have more varied perceptions. These
findings suggest a potential link between educational attainment and
experiences or attitudes toward corruption, particularly among those
with secondary-level education.

```{r}
#| label: fig-age-versus-corruption
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Age Group Vs. Corruption Perception. This faceted bar chart plots the relationship between the age group of respondents and their different corruption perception levels in Nigeria."

# Create the faceted bar chart
ggplot(cleaned_data, aes(x = age_group, fill = age_group)) +
  geom_bar() +
  facet_wrap(~corruption_presidency) +
  labs(
    title = "Corruption Perceptions across Age Groups",
    x = "Age Group",
    y = "Count of Respondents",
    fill = "Age Group"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

```

@fig-age-versus-corruption shows the relationship between respondents'
age group which was obtained by asking the respondent *How old are you?* and their perceptions of corruption in Nigeria. Each panel corresponds to one of four corruption perception levels. Within
each panel, the bars represent the count of respondent across five age
groups (18-24, 25-34, 35-44, 45-54, and 55+), enabling a clear
comparison of how perceptions vary by age.

The 25–34 and 35–44 age groups have the highest counts in the High
Corruption and Some Corruption categories, suggesting that middle-aged
respondents are more likely to perceive corruption as significant. In
contrast, younger respondents (18–24) and older respondents (55+) report
lower levels of corruption perception in these categories.In the All Corrupt category, the 25–34, 35–44, and 55+ age groups dominate, while the 18–24 and 45–54 groups have lower counts. The No
Corruption category has low counts across all age groups, indicating
that very few respondents, regardless of age, perceive the absence of
corruption.

These results indicate that middle-aged respondents (25–44) are the most
likely to report high levels of corruption perception, while younger and
older groups report less. The consistently low counts in the No
Corruption category reflect a widespread belief that corruption is a
persistent issue in Nigeria.

```{r}
#| label: fig-ruralurban-versus-corruption
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Rural/Urban Vs. Corruption Perception. This stacked bar chart visualises the relationship between respondents' corruption perceptions and weather they live in rural or urban areas in Nigeria."

# Summarize the data
corruption_vs_urban <- cleaned_data %>%
  count(urban_rural, corruption_presidency) %>%
  group_by(urban_rural) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup()

# Create the stacked bar chart
ggplot(corruption_vs_urban, aes(x = urban_rural, y = proportion, fill = corruption_presidency)) +
  geom_bar(stat = "identity", position = "fill") +
  labs(
    title = "Relationship between Rural vs. Urban Respondents and Corruption Perceptions",
    x = "Respondent Location",
    y = "Proportion of Respondents",
    fill = "Corruption Perception"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

# Convert corruption_presidency to numeric 
cleaned_data$corruption_numeric <- as.numeric(factor(cleaned_data$corruption_presidency))

```

The relationship between respondents' residential location (rural vs.
urban) and their perceptions of corruption is shown in
@fig-ruralurban-versus-corruption Each bar represents the proportion of
respondents within rural and urban groups who hold different corruption
perceptions.

The chart shows that rural and urban respondents share similar
patterns of corruption perception. The majority of respondents in both
groups fall under the Some Corruption and High Corruption categories,
with Some Corruption being slightly more prevalent. The All Corrupt
category is the third most common perception in both groups, while the
No Corruption category is consistently the least represented.

Residential location does not significantly influence corruption perceptions, as
rural and urban respondents report similar distribution patterns across all 
categories. The results suggest that perceptions of corruption are pervasive and
relatively consistent across geographic locations.

```{r}
#| label: fig-gender-versus-corruption
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Gender Vs corruption perceptions. This faceted bar chart shows the relationship between a respondents' gender and how they perceive corruption within the presidency in Nigeria."

# Summarize the data
corruption_vs_gender <- cleaned_data %>%
  count(gender, corruption_presidency) %>%
  group_by(gender) %>%
  mutate(proportion = n / sum(n)) %>%
  ungroup()

# Create the faceted bar chart
ggplot(cleaned_data, aes(x = corruption_presidency, fill = corruption_presidency)) +
  geom_bar() +
  facet_wrap(~gender) +
  labs(
    title = "Corruption Perceptions by Gender",
    x = "Corruption Perception",
    y = "Count of Respondents",
    fill = "Corruption Perception"
  ) +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

@fig-gender-versus-corruption shows the the relationship between
respondents' gender (male and female) and their perceptions of
corruption. Each panel represents one gender, displaying the count of
respondents across four corruption perception categories: All Corrupt,
High Corruption, Some Corruption, and No Corruption.

The chart shows that both male and female respondents share similar
patterns of corruption perception. The High Corruption and Some
Corruption categories have the highest counts for both genders,
indicating these are the most commonly held views. The All Corrupt
category has fewer respondents, while the No Corruption category
consistently has the lowest counts for both genders.

# Model {#sec-model}

## Model set-up

A Bayesian ordinal logistic regression model is used to estimate the
relationship between various demographic factors and perceptions of
corruption in the presidency. Ordinal logistic regression is a
statistical technique used for ordered categorical outcomes to predict
the cumulative probability of being in a particular category or below,
given a set of predictor variables.

My model includes four independent demographic variables: `education`,
`gender`, `urban_rural`, and `age_group`. The dependent variable is
`corruption_presidency`, which has four ordered categories: "None,"
"Some," "Most," and "All."

The ordinal logistic regression model I used is: \begin{equation}
\log\left(\frac{P(y_i \leq j)}{P(y_i > j)}\right) = \alpha_j - \left(\beta_1 \times \text{education} + \beta_2 \times \text{gender} + \beta_3 \times \text{urban rural} + \beta_4 \times \text{age group}\right)
\end{equation}

$$
\begin{aligned}
\beta_1 & \sim \mbox{Normal}(0, 2.5)\\
\beta_2 & \sim \mbox{Normal}(0, 2.5)\\
\beta_3 & \sim \mbox{Normal}(0, 2.5)\\
\beta_4 & \sim \mbox{Normal}(0, 2.5)\\
\alpha_j & \sim \mbox{Normal}(0, 2.5) \quad \text{for } j = 1, 2, 3\\
\pi_{ij} & \sim \mbox{Dirichlet}(\alpha = 1)
\end{aligned}
$$ where:

-   $y_i$ represents the ordinal outcome for respondent $i$, indicating
    their perception of corruption in the presidency.

-   $P(y_i \leq j)$ is the cumulative probability of respondent $i$
    falling into category $j$ or below.

-   $\alpha_j$ are the threshold parameters that define the boundaries
    between the corruption categories.

-   $\beta_1, \beta_2, \beta_3, \beta_4$ are the coefficients
    corresponding to the predictors:

    -   $\beta_1$: Education level of the respondent.
    -   $\beta_2$: Gender of the respondent (male or female).
    -   $\beta_3$: Urban or rural location of the respondent.
    -   $\beta_4$: Age group of the respondent.

-   $\pi_{ij}$ is the probability that respondent $i$ falls into
    corruption category $j$.

In this model, the priors reflect a belief that the predictors explain
about 20% of the variability in the outcome, as defined by the
$R^2(0.2)$ prior. The Normal priors with a mean of 0 and standard
deviation of 2.5 for the coefficients and threshold parameters are
relatively weakly informative, ensuring the data have a significant role
in determining the posterior distributions.

The threshold parameters ($\alpha_j$) separate the ordered categories of
the outcome variable, while the Dirichlet prior for $\pi_{ij}$ assumes
no strong prior preference for any specific corruption category.

The use of these priors provides flexibility for the model while
preventing overfitting. The posterior distributions for the coefficients
and thresholds, along with diagnostic checks for model convergence and
predictive accuracy, can be found in the Appendix.

### Model Assumptions

The model relies on the following assumptions:

-   Proportional Odds: The relationship between the predictors and the
    log-odds of being in a higher corruption category is constant across
    thresholds.
-   Ordinal Outcomes: The response categories are ordered, but the
    distances between them are not quantitatively defined.
-   Independence of observations: Responses from different individuals
    are assumed to be independent.

## Model Justification

The Bayesian approach was chosen because of its flexibility in
incorporating prior beliefs and its ability to provide real
quantification for uncertainty for parameter estimates. An ordinal
logistic regression model is appropriate because the outcome variable
(perceived corruption) is inherently ordered.

The predictors-education, gender, urban/rural location, and age group-
were selected based on their theoretical relevance to perceptions of
governmental corruption in Nigeria, as discussed in the data section.
For example, urban and rural individuals may influence exposure to
governance systems, while education might shape perceptions of systemic
corruption.

The modeling decisions align closely with the structure and
characteristics of the dataset to ensure interpretability and relevance.
Age was included as a categorical variable reflecting predefined groups,
allowing the model to capture non-linear relationships between age and
perceptions of corruption. Education was treated as a categorical
variable to account for differences across distinct milestones, such as
"Primary school completed" and "Post-graduate qualifications." Gender
was modeled as a binary variable ("Male" or "Female") to reflect
societal influences on corruption perceptions. Urban versus rural
context was included as a binary variable to capture potential
differences in exposure to governance and information access. While the
dataset includes regional information, province effects were excluded to
maintain focus on demographic factors and simplify the model. Finally,
the ordinal logistic regression framework was chosen to respect the
ordered nature of the outcome variable ("None," "Some," "Most," "All")
without assuming equal distances between categories. These decisions
ensure the model is both consistent with the data and well-suited to
addressing the research question.

# Results {#sec-results}

## Corruption Perceptions Model

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Summary of the Ordinal Logistic regression model"
#| warning: false

library(broom.mixed)
library(kableExtra)

# Extract summary
model_summary <- broom.mixed::tidy(bayesian_model)

# Render as a table
knitr::kable(
  model_summary,
  caption = "Summary of the Bayesian Model",
  digits = 2
) %>%
  kableExtra::kable_styling()

```

The analysis revealed distinct trends in Corruption perceptions in
different demographics. Our results are summarized in @tbl-modelresults. 
This presents the estimated coefficients and standard
errors for each predictor included in the model. The dependent variable
is the perception of corruption in the presidency, categorized into
ordered levels. The table includes predictors such as education level,
gender, urban or rural location, and agre group, which are modeled to
explain variations in corruption perceptions. Each predictor's
coefficient represents its effect on the log odds of perceiving
corruption in the presidency at a specific level relative to lower
levels, holding other variables constant.

Higher education levels(e.g., university or post-graduate
qualifications) arre associated with positive effects on corruption
perceptions, indicating a greater likelihood of perceived higher levels
of corruption. Begin female shows a small positive association with
higher corruption perceptions. Residing in rural areas is negatively
associated with corruption perceptions, suggesting lower perceptions of
corruption compared to urban counterparts. Agegroups show minimal
variation in their association with corruption perceptions, with no
significant deviations observed among different age brackets.

## Predicted Probabilities

```{r}
#| echo: false
#| eval: true
#| label: fig-predicted-probabilities
#| fig-cap: "Predicted probabilities of high corruption perceptions by key predictors."

library(ggplot2)

# Simulating or using your posterior data
predicted_data <- data.frame(
  Predictor = names(coef(bayesian_model)),
  Estimate = coef(bayesian_model)
)

# Plot with adjusted x-axis labels
ggplot(predicted_data, aes(x = Predictor, y = Estimate)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Rotate and reduce size
    axis.title.x = element_text(size = 10)
  ) +
  labs(
    title = "Predicted Probabilities by Demographics",
    x = "Predictor",
    y = "Probability"
  )

```

@fig-predicted-probabilities illustrates the predicted probabilities
derived from the Bayesian ordinal logistic regression model for
different demographic predictors influencing corruption perceptions in
Nigeria. Each bar represents the estimated impact of a specific
predictor on the likelihood of perceiving higher corruption levels in
the presidency.

Higher education levels consistently show positive probabilities,
indicating that individuals with higher education levels are more
likely to perceive higher corruption levels. Living in rural areas has a
significantly negative probability, suggesting that rural residents are
less likely to perceive high corruption compared to urban residents.
Gender shows a moderate positive probability, with females slightly more
likely to perceive higher corruption levels than males. There seems to
be minimal variation observed across age groups, suggesting that age
does not substantially influence perceptions of corruption.

The results show that higher levels of education, particularly
university completion and post-secondary qualifications, are associated
with greater perceptions of corruption in the presidency. Urban
residents demonstrate significantly higher perceived corruption levels
compared to rural residents. Gender differences reveal that female
respondents show slightly higher corruption perceptions, while age
groups show minimal variation. These findings are shown in the
regression estimates and predicted probabilities, as visualized in
@fig-predicted-probabilities and summarized in @tbl-modelresults.

# Discussion {#sec-discussion}

This paper looks into the demographic disparities of corruption
perception in the Nigerian presidency, using survey data to examine
variables like education, gender, age and urban versus rural residence.
By looking at these factors, the study aims to inform targeted
anti-corruption camapigns and policy making. The Bayesian modeling
approach helps us understand the relationship between corruption
perceptions and these demographic variables.

## Tareting Education Levels in Anti-Corruption Campaigns

The relationship between education and corruption perceptions shows that
respondents with secondary school education are the most likely to
perceive high levels of corruption in the presidency. This supports
broader research showing that moderate levels of education are linked to
increased political awareness and accountability demands, as individuals
in this group are better equipped to identify systemic corruption but
may lack the structural access to address it effectively
[@transparency2019]. Individuals with higher education, such as
university graduates, report more moderate corruption perceptions, this
might largely beceause of their exposure to diverse viewpoints and
greater access to institutional frameworks that allow for political
skepticism. These suggest that anti-corruption campaigns targeting this
moderately educated demographic should focus on actionable reforms and
accountability mechanisms to harness their heightened awareness
productively [@u4nigeria2023].

## Urban-Rural Dynamics in Perception

Urban respondents are somewhat more likely than their rural counterparts
to perceive high corruption in the presidency. This is consistent with
findings that urban populations have greater access to investigative
reporting and public discourse on corruption, which can amplify
awareness and perceptions [@u4nigeria2023]. On the other hand, rural
respondents, who may rely on localized patronage networks for essential
services, often exhibit more neutral or unclear perceptions of
corruption, influenced by limited media access and dependency on these
informal systems. Campaigns should focus on leveraging urban platforms
like social media, community-based programs, and public dialogues to
engage important urban audiences, while also expanding rural outreach
through local radio, grassroots organizations, and educational
interventions tailored to these settings.

## The Role of Age, Youth & Gender Engagement

Younger respondents, particularly those aged 25–34, consistently report
higher perceptions of corruption compared to older age groups. This
aligns with existing research that highlights younger demographics as
more vocal in their demands for systemic change in a country like
Nigeria, where over 60% of the population is under 25 [@UNFPA2023],
targeted anti-corruption initiatives that harness youth energy are
important. These efforts should include digital campaigns on platforms
frequently used by younger audiences, interactive town hall discussions,
and youth-led advocacy initiatives that focus on transparency and
accountability. Importantly, such campaigns need to have clear pathways
for action, making it easy for youth to channel their frustration into
productive engagement with governance structures.

While gender differences in corruption perceptions were less pronounced
in our paper, prior research shows that women often experience
corruption indirectly, particularly in accessing healthcare or education
systems [@GenderCorruption2021]. This shows the need for anti-corruption
campaigns that address systemic barriers disproportionately impacting
women, emphasizing institutional reforms in sectors that are important
to gender equity. Moreover, trust in institutions plays an important
role in shaping perceptions. Corruption scandals and perceived impunity
have a way of eroding trust, particularly in urban areas where political
engagement is higher[@afrobarometer2007]. Restoring institutional
credibility through transparency and consistent enforcement of
anti-corruption measures can increase the effectiveness of campaigns
targeting both genders.

## Policy Implications

This paper is focused on looking at trends in the corruption perception
between different demographics and seeing how these differences can be
used to maximize anti corruption campaigns and policy in
Nigeria.From this paper, we find different actionable strategies for
improving the effectiveness of anti-corruption campaigns in Nigeria.
First, targeting rural and urban populations differently might be
effective. Urban residents, who have greater access to investigative
media and public debates, are more likely to perceive high corruption.
Campaigns in urban areas should therefore focus on digital platforms
like social media and mobile apps, as well as partnerships with
investigative journalists and public accountability platforms to engage
this group effectively. In rural areas, where reliance on patronage
systems and limited media access shape perceptions, grassroots
communication strategies like community radio, local town halls, and
collaborations with trusted community leaders are better suited to
disseminate anti-corruption messaging.

Education also emerges as an important determinant of corruption perceptions.
Secondary school graduates exhibit the highest levels of awareness, this
suggests that civic education campaigns should focus on this demographic
to harness their increased engagement. Integrating anti-corruption
education into school curricula could early awareness and long-term
attitudinal changes.Youth engagement is very important, given that
younger age groups, particularly those aged 25–34, report the highest
levels of dissatisfaction with corruption. Social media campaigns
tailored to platforms popular with youth, together with interactive town
hall discussions and youth-led grassroots initiatives, could channel
this demographic's energy into organized anti-corruption efforts.

Lastly, rebuilding public trust in institutions should be a priority.
Transparent hiring practices, visible enforcement of anti-corruption
laws, and strengthened whistle blower protections are important steps to
showing institutional accountability. Localized outreach through
community leaders can also help fix the trust divide, particularly in
rural settings. Policymakers should put these findings into a
data-driven framework so they can adapt strategies based on demographic
trends and feedback.

## Weaknesses and next steps

While this paper tells us a lot about the demographic distributions of
corruption perception about the presidency in Nigeria, there are a few
limitations that have to be considered. The survey data relies on
self-reported perception, which can be influenced by social desirability
bias which refers to the situation where respondents' endorse more
favorable responses in order to enhance their own
self-presentation[@bias2022].For example, high-profile corruption
scandals or electoral promises happening close to survey periods may
distort responses. The data does not account for regional differences
within these urban rural classifications, which could change the
understanding of corruption perceptions further.The last limitation is
with the bayesian model, because while the Bayesian model accounts for
uncertainty, it does not address potential unobserved variables, like
political affiliations or socioeconomic status, which may affect the
relationships we say.

Future research should look more into longitudinal trends in corruption
perceptions to look at how demographic factors evolve over time. This
could be achieve through tracking respondents across electoral cycles or
policy interventions to observe shifts in attitudes. Additionally,
looking more into regional subcategories could show us some localized
patterns, particularly in states that have different governing
structures and resource allocation systems. Expanding the dataset to
include qualitative information from interviews and focus groups would
also give more information about how demographic characteristics relate
to peoples' lived corruption experiences.

\newpage

\appendix

# Appendix {.unnumbered}

# Surveys, Sampling, and Observational Data

## Aforbarometer Methodology Overview and Evaluation {#sec-survey-appendix}

### Overview

Afrobarometer is a  pan-African, non-partisan survey research network established to measure public attitudes on democracy, governance, economic reform, civil society, and quality of like in Africa. It relies predominantly on in-person interviews using a standard questionnaire. The project covers up to 40 countries in Africa. The information collected is intended to be used by various parties including policymakers, activists,teachers and scholars. The surveys are conducted using a standardized questionnaire tailored to specific countries through a process called "indigenization", ensuring relevance to local contexts while keeping comparability across nations [@survey_manual_2022].

### Population and sampling frame

The Afrobarometer surveys target all citizens of voting age, typically 18 and above, excluding individuals living in institutions such as prisons, military barracks, or hospitals. The sampling frame is based on the most recent official national census. When census data is outdated, population estimates or projections are used to ensure relevance. The surveys employ a clustered, stratified, multi-stage probability sample design, where:

- Clustered sampling groups respondents geographically to reduce costs and logistical complexity.

- Stratified sampling divides the population into subgroups (e.g., urban/rural or regions) to ensure proportional representation.

- Multi-stage sampling involves multiple rounds of random selection, narrowing down from regions to individual households and respondents.

### Respondent Recruitment

Afrobarometer follows strict protocols to ensure randomness and gender balance during respondent selection. Interviewers alternate genders with each new interview, starting each day with the gender opposite to the last interview conducted the previous day. If no eligible respondent of the next household. Interviews are conducted only after securing informed consent from the respondents, making sure to keep up with ethical guidelines to ensure respect and confidentiality [@survey_manual_2022].

### Smpling Approach

The Probability Proportional to Population Size (PPPS) sampling method makes sure that regions with larger populations have a higher chance of being selected. In urban areas, the sampling process typically has four stages: selection of Primary Sampling Units (PSUs) based on population size, random selection of Enumeration Areas (EAs) within PSUs, systemic selection of households within each EA, random selection of respondents within household [@survey_manual_2022]. In rural areas, an additional stage may be introduced to make sure representativeness across dispersed populations. By using PPPS and randomization at every stage, Afrobarometer minimizes biases and makes sure data is collected efficiently [@survey_manual_2022].

### Strengths and Limitations

**Strengths:**

- Random Selection: Ensures every adult citizen has an equal chance of being selected, enhancing representativeness.

- Proportional Sampling: PPPS ensures larger and more populous regions are adequately represented.

- Localized Surveys: Conducted in local languages to ensure accessibility and reduce misinterpretation of questions.

**Limitations:**

- Institutional Exclusion: Citizens in prisons, military barracks, and similar institutions are excluded, potentially overlooking significant perspectives.

- Conflict Zones: Areas affected by conflict or logistical challenges may be excluded, limiting geographic coverage.

- Africa-Wide Generalization: While Afrobarometer provides national-level information, its findings cannot be generalized to the entire continent due to contextual variations.

## Ethics considerations {#sec-ethics}

### Overview of Afrobarometer’s Ethical Framework

Afrobarometer adheres to  ethical standards to ensure the protection and dignity of survey respondents. Here are the important components of their ethical framework:

- **Informed Consent**: Respondents are provided with detailed information about the survey's purpose, topics, and their rights before participation. Verbal consent is obtained to account for literacy variations across surveyed populations [@survey_manual_2022].
- **Confidentiality and Anonymity**: Personal identifiers are removed during data processing, and data are stored securely to maintain respondent privacy.
- **Cultural Sensitivity**: Questionnaires are carefully pre-tested and translated into local languages, ensuring they are culturally appropriate and easily understood.

These practices align with established ethical principles in social science research, including those outlined by the World Association for Public Opinion Research [@waporethics].

### Ethical Challenges in Corruption Perception Surveys

Researching perceptions of corruption has unique ethical challenges because of the sensitive and sometimes dangerous nature of corruption:

- **Risks to Respondents**: Respondents may fear repercussions for expressing negative views about governance. Afrobarometer mitigates this risk by emphasizing anonymity and employing well-trained enumerators who follow strict protocols [@survey_manual_2022].
- **Power Dynamics**: Interviewers may inadvertently influence responses due to perceived power imbalances. Afrobarometer addresses this by alternating gender in interviews and training enumerators to minimize bias.

### Sampling and Recruitment Ethics

Afrobarometer’s methodology includes a very detailed sampling framework designed to enhance inclusivity and fairness. However, there are still some ethical concerns:

- **Exclusion of Institutional Populations**: Individuals in prisons, military barracks, and similar institutions are excluded, potentially omitting perspectives from disenfranchised groups [@survey_manual_2022].
- **Gender Quotas**: The enforcement of equal gender representation ensures that women's voices are equally captured, addressing historical disparities in research participation [@survey_manual_2022].
- **Regional and Security Constraints**: Some areas, particularly conflict zones, may be inaccessible, leading to under representation in survey findings.This is particularly the cause in North-Eastern Nigeria which as a result of criminal groups like Boko Haram, has made it virtually impossible to access that region safely.

### Ethical Data Management and Reporting

Afrobarometer takes significant steps to ensure ethical data handling and dissemination:

- **Data Security**: Survey data are encrypted during storage and transmission, ensuring compliance with data protection regulations.
- **Transparent Reporting**: Reports explicitly document survey limitations, including regions excluded due to security or logistical challenges.
- **Avoiding Harm**: Results are contextualized to prevent misinterpretation or harm to surveyed populations. Findings are shared responsibly, avoiding language that could inflame tensions.

### Limitations and Future Ethical Enhancements

Despite its robust framework, Afrobarometer’s approach could be improved in the following areas:

- **Inclusion of Marginalized Populations**: Future surveys could include anonymous digital surveys to take from institutionalized or transient populations.
- **Addressing Non-Response Bias**: Enhanced recruitment strategies, such as partnerships with community organizations, could increase participation among underrepresented groups.
- **Community Engagement**: Sharing results directly with communities through accessible formats (e.g., infographics or local town halls) would increase transparency and accountability.

### Conclusion

The ethical considerations underpinning Afrobarometer’s methodology ensure the credibility and validity of its findings while safeguarding respondent dignity. However, addressing remaining gaps—such as the exclusion of certain populations and challenges in conflict zones—would further enhance the ethical rigor of corruption perception research in Africa. By adhering to established ethical frameworks and remaining responsive to emerging challenges, studies like this can continue to give important information while maintaining the highest ethical standards.To do my part in ensuring ethical standards are upheld, i have included a datasheet for the Afrobarometer dataset.

# Additional data details

@fig-demo The figure provides an overview of demographic variables and corruption perceptions from the Afrobarometer Nigeria Round 9 survey. Rural residents constitute a larger share of the sample, aligning with Nigeria's predominantly rural demographic landscape. Corruption perception data shows that "High Corruption" and "Some Corruption" dominate public opinion. The education level distribution highlights the prevalence of basic education, with the majority of respondents having completed secondary school. Finally, the age distribution indicates that the 25–34 age group is the most represented cohort, suggesting that young adults play a crucial role in shaping perceptions of corruption and governance in Nigeria.
![Demographic Distributions of the Afrobarometer Data](combined_plot.png){#fig-demo width="120%"}

# Model details {#sec-model-details}

## Posterior Predictive Check

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheck
#| fig-cap: "Posterior Predictive Check for the Bayesian ordinal logistic regression model."

# Load the Bayesian model
bayesian_model <- readRDS(file = here::here("models/bayesian_model.rds"))

# Posterior Predictive Check
pp_check(bayesian_model) +
  theme_classic() +
  labs(
    title = "Posterior Predictive Check",
    x = "Observed Data",
    y = "Density"
  ) + 
  scale_x_continuous(
    name = "Observed Data",     
    breaks = seq(0, 1, by = 0.1), 
    limits = c(1.1, 2.8)             
  )

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-posteriorvsprior-education
#| fig-cap: "Posterior vs Prior Comparison for Education Parameters."

# Education Parameters
education_params <- c(
  "educationInformal schooling only",
  "educationSome primary schooling",
  "educationPrimary school completed",
  "educationSome secondary schooling",
  "educationSecondary school completed",
  "educationPost-secondary qualifications",
  "educationSome university",
  "educationUniversity completed",
  "educationPost-graduate"
)

posterior_vs_prior(bayesian_model, pars = education_params, group_by_parameter = TRUE) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(
    title = "Posterior vs Prior Comparison (Education Parameters)",
    x = "Parameter Value",
    y = "Density",
    color = "Distribution"
  )

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-posteriorvsprior-demographic
#| fig-cap: "Posterior vs Prior Comparison for Demographic Parameters."

# Demographic Parameters
demographic_params <- c(
  "genderFemale",
  "urban_ruralRural",
  "age_group25-34",
  "age_group35-44",
  "age_group45-54",
  "age_group55+"
)

posterior_vs_prior(bayesian_model, pars = demographic_params, group_by_parameter = TRUE) +
  theme_minimal() +
  scale_color_brewer(palette = "Set2") +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(
    title = "Posterior vs Prior Comparison (Demographic Parameters)",
    x = "Parameter Value",
    y = "Density",
    color = "Distribution"
  )

```

## Diagnostics

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-diagnostic-traceplot
#| fig-cap: "Trace plot for selected parameters in the Bayesian ordinal logistic regression model."

# Trace Plot
mcmc_trace(as.array(bayesian_model), pars = c("educationSome primary schooling", "urban_ruralRural")) +
  theme_minimal() +
  labs(
    title = "Trace Plot for Selected Parameters",
    x = "Iteration",
    y = "Parameter Value"
  )

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-diagnostic-rhat
#| fig-cap: "Rhat values for model convergence in the Bayesian ordinal logistic regression model."

# Rhat Plot
mcmc_rhat(rhat(bayesian_model)) +
  theme_minimal() +
  labs(
    title = "Rhat Plot",
    x = "Rhat Values",
    y = "Parameters"
  )

```

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-diagnostic-ess
#| fig-cap: "Effective sample size plot for parameters in the Bayesian ordinal logistic regression model."

# Effective Sample Size Plot
mcmc_neff(neff_ratio(bayesian_model)) +
  theme_minimal() +
  labs(
    title = "Effective Sample Size Plot",
    x = "Effective Sample Size Ratio",
    y = "Parameters"
  )

```

\newpage

# References